---
output:
  pdf_document: default
  html_document: default
---

MICE IMPUTATIONS FOR MISSING DATA IN R

Team Members - ME18B148, ME18B135, ME18B136, ME18B111, ME18B162, ME18B170

The following is the R code used for MICE imputations for the given dataset.

First importing the given dataset and respective labels.
```{r}
library(readxl)
train_x <- read.csv("D:/Sem 5/IDA project dataset/train_x.csv", na.strings = c(""," ","NA"))
train_y <- read.csv("D:/Sem 5/IDA project dataset/train_y.csv", na.strings = c(""," ","NA"))

```

Now merging both of these into one dataframe.
```{r}
data <- train_x
data$Label <- train_y$Label
```

Let us plot the missingness map to interpret the missing data. Amelia package is generally used to plot the missingness map.
```{r}
library(Amelia)
missmap(data[-1], col = c('grey', 'steelblue'), y.cex=0.5, x.cex=0.8)
```

Now let us use the mice package to impute the missing data. We have used pmm method for Numeric Variables and cart method for Categorical Variables.
```{r}
library(mice)
imp.data <- mice(data, m=1, method = c("", "pmm", "pmm", "cart", "cart", "cart", "pmm", "pmm", "pmm", "pmm", "pmm", "cart"), printFlag = FALSE)
```

Storing the imputed data.
```{r}
data <- complete(imp.data)
```

In the previous line of code, all features were not imputed, this might be a restriction of mice package. So we had to impute the 'Expense' column soley in the below lines of code.
```{r}
imp.data2 <- mice(data, m=1, method = c("", "pmm", "", "", "", "", "", "", "", "", "", ""), printFlag = FALSE)
```

Finally storing the completely imputed data.
```{r}
data <- complete(imp.data2)
```

Let us check for any missing data after imputations.
```{r}
sort(sapply(data, function(x) {sum(is.na(x))}), decreasing = TRUE)
```
As you can see there is no data missing and all missing data has been imputed.

Plotting Feature vs Feature to interpret imputed data vs existing data. The imputed data is almost similar to the existing data, So we can say that MICE imputations worked well.
```{r}
xyplot(imp.data, Score1 ~ Score2)
xyplot(imp.data, Score1 ~ Score3)
xyplot(imp.data, Score1 ~ Score4)
xyplot(imp.data, Score1 ~ Score5)
xyplot(imp.data2, Expense ~ Income)
```

Splitting the data into X and Y sets.
```{r}
train_y$Label <- data$Label
train_x <- data[, 1:11]
```

Finally exporting the data to csv files, and then trained using python.
```{r}
write.csv(train_x, "D:/Sem 5/IDA project dataset/train_x_imputed.csv", row.names = FALSE )
write.csv(train_y, "D:/Sem 5/IDA project dataset/train_y_imputed.csv", row.names = FALSE )

```

